### 2016-12-12 paper/paper.tex
Starting work on the paper.

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\usepackage{listings,lstlangcoq}

\urlstyle{sf}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{COQPL '16}{January 23, 2015, St. Petersburg, Florida, United States} 
\copyrightyear{2016} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Certified Desugaring of Javascript Programs using Coq}
\subtitle{Experience Report}

\authorinfo{Marek Materzok}
           {INRIA, University of Wroc\l{}aw}
           {marek.materzok@inria.fr}

\maketitle

\newcommand{\lambdajs}{$\lambda_\textrm{JS}$}

\begin{abstract}
JavaScript is a programming language originally developed for client-side
scripting in Web browsers; its use evolved from simple scripts to
complex Web applications. It has also found use in mobile applications,
server-side network programming, and databases.

A number of semantics were developed for the JavaScript language.
We are specifically interested in two of them: JSCert and \lambdajs{}.
In order to increase our confidence that the two semantics correctly
model JavaScript, we try to relate them formally using Coq. The size and complexity
of the two semantics makes this a complex problem with many obstacles
to be overcome.
\end{abstract}

\category{D.3.1}{Programming Languages}{Formal Definitions and Theory}

% general terms are not compulsory anymore, 
% you may leave them out
%\terms
%term1, term2

\keywords
Coq, JavaScript, ECMAScript, mechanized semantics

\section{Introduction}

The JavaScript language is standardized by ECMA under the name
ECMAScript. Several versions of the standard now exist, the latest
two being ECMAScript 5.1 (ES5, released in 2011) and ECMAScript 6
(ES6, released in 2015). Its role is to make portability between different
JavaScript implementations possible by specifying the intended behavior
of the various language constructs and built-in functions of JavaScript.
The specification also makes it possible to develop formal semantics
for the language, which can then be used for, e.g., developing
static analysis tools.

One such semantics is JSCert~\cite{Bodin-al:POPL14}, which has the
goal of formalizing the ECMAScript (version 5.1) specification in the form of
a pretty-big-step semantics~\cite{Chargueraud:ESOP13} expressed in 
the form of a Coq inductive predicate. It aims to be as close to the
(informal English) specification as possible by maintaining close correspondence
between the evaluation rules and steps in the ES5 pseudocode. The project
also includes JSRef, a reference interpreter for JavaScript written in Coq
and proven correct with respect to JSCert. The interpreter is also tested
using the official ECMA test suite, test262. Thanks to the correctness proof,
testing of JSRef increases the confidence we can have that JSCert
correctly models ES5.

Another semantics we are interested in is \lambdajs. It aims to capture
the semantics of JavaScript by defining a simple core language
(based on the untyped call-by-value lambda calculus), and then
transforming (or ``desugaring'') the full JavaScript language into this
simple core. The \lambdajs{} semantics was originally developed to model the 
ECMAScript 3 specification~\cite{Guha-al:ECOOP10}, and it was later
revised~\cite{Politz-al:DLS12} to handle ECMAScript 5.1
features\footnote{The authors named the revised language S5; however,
we continue to name it \lambdajs{} for simplicity.}.
The interpreter for the core language and the JavaScript-to-core
transformation were developed using 
OCaml\footnote{\url{https://github.com/brownplt/LambdaS5}};
the interpreter was tested using the official test262 test suite.
The \lambdajs{} semantics could be useful for tool developers, as it
allows the tools to analyze JavaScript programs by working
on the simple core language of \lambdajs, having all the complexities
of JavaScript handled in the desugared code. It was already
successfully used to verify AdSafe. %TODO citation

We believe is worthwhile to formally relate the two semantics. One reason is,
because the two semantics use a different approach, relating
them will increase the confidence we can place in both of them,
and is likely to find inconsistencies or other problems. Another reason is
that having \lambdajs{} formalized and certified (by proving correctness
with respects to JSCert) enables development of certified analysis tools
using \lambdajs{} as the core.

This is easier said than done. The task is very complex, mostly due
to the complexities of the JavaScript language. The ES5 specification,
upon which JSCert and \lambdajs{} are based, is 258 pages long and 
has 16 chapters. The JSCert pretty-big-step semantics, which models
the specification, has over 900 rules; the entire Coq source code
for the semantics has over 10000 lines. The \lambdajs{} environment
file, which contains the ECMAScript algorithms expressed in core
\lambdajs{} code, is almost 5000 lines long.

To reduce the complexity of the task, we are currently working
only on proving \lambdajs{} correct with respect to JSCert,
i.e., that if the \lambdajs{} program successfully terminates,
then its result matches the result given by JSCert. We are also
restricting our attention to ECMAScript strict mode, proving the
correctness of non-strict features only when it is convenient.

We were able to prove correctness of a large part of the
JavaScript semantics, including function calls, lexical environment
handling, and most operations involving objects. The work is still
in progress to prove the remaining parts, which include integer
operations, the
defineOwnProperty method, and elements of the standard library.

The development is publicly available and can be downloaded from GitHub:

\url{https://github.com/tilk/LambdaCert}

\section{Formalization of \lambdajs{}}

For the purpose of relating the two semantics, we first
developed a formalization of the \lambdajs{} core language.
This was necessary because no such formalization existed previously;
a formalization of the original, ES3-based \lambdajs{}
was available\footnote{\url{https://github.com/brownplt/LambdaJS}},
but the two versions of \lambdajs{} differ too much for this
to be useful.

We decided to formalize the \lambdajs{} core using a
pretty-big-step semantics, as in JSCert, to make the job of
relating the two semantics easier. We have also developed
an interpreter for \lambdajs{}, which we proved sound and
complete with respect to the pretty-big-step semantics.
The proof script was developed using domain-specific tactics,
which made it simple and robust.

At this stage we stumbled upon an interesting performance issue 
involving the Coq proof assistant. The proof of completeness required
inverting the pretty-big-step semantics judgment; this turned out
to be very slow, and also generated large proof terms, which increased
the memory footprint. The problem is well-known~\cite{Monin:Coq2}; 
our solution involved
precomputing the required inversion principles using
\texttt{Derive Inversion}, and using the precomputed inversion
principles to create a specialized inversion tactic.

We have also implemented in Coq the other two crucial components
of the \lambdajs{} semantics: the desugaring function and the 
environment. The desugaring was implemented as Coq functions,
taking JavaScript statements and expressions (as defined in
JSCert) to \lambdajs{} expressions. The environment, which
implements the complexities of the JavaScript semantics
in \lambdajs{} code, was translated from \lambdajs{} syntax
to Coq definitions with the help of an OCaml parser and
our interpreter. The resulting Coq file is over 15000 lines long
and takes almost a megabyte of disk space!

\section{Correctness proof for \lambdajs{} desugaring}

In order to establish correctness, it was first needed to create
a set of predicates specifying the relationship between 
JavaScript and \lambdajs{} language features. As \lambdajs{} was
designed to implement JavaScript, the relationship is fairly
simple and natural. One place where the two languages differ
is the object heap, which is used in \lambdajs{} to implement
several JavaScript language features, including the
JavaScript lexical environment. To manage this, we use a bisimulation
relation, which relates JavaScript objects and environment records
to \lambdajs{} objects.

For the proof, we also need to keep track of certain invariants,
which state that, e.g., the JavaScript objects and their
\lambdajs{} representations are correctly formed. 
The invariants are split into two relations, where one
involves the evaluation contexts, and the other -- the heaps.
Both of these involve the bisimulation relation, but the
context invariant uses it only positively. As the bisimulation
relation only grows during the evaluation (as new objects
are created), this makes it possible to keep the proofs simple.

The correctness proof is structured using well-founded 
induction on the depth of the \lambdajs{} pretty-big-step
semantics derivation. This grants a lot of freedom in structuring
the proof, which is important because of certain structural
differences between JSCert and \lambdajs{}, which involve,
e.g., handling of ES5 references. We use three kinds of induction
hypotheses: for JavaScript statements, expressions and
function calls. The separate induction hypothesis for function
calls was required because of the complexity of calling functions
in JavaScript, and also because function calls can happen in
many possible places in JavaScript, including every
object to primitive conversion.

We prove many lemmas about the different language constructs
of JavaScript, including the internal ``specification functions''
(e.g., type conversions, operations on objects and environment
records, etc.). Typically, the proof for a lemma consists of
a series of forward reasoning steps and case analyses, and ends with
constructing a JSCert semantics derivation and proving that
the invariants were preserved. The forward reasoning is largely
automated using domain-specific tactics; the JSCert derivation
is constructed using the \texttt{eauto} tactic with a specialized
hint database (\texttt{nocore} is used to avoid using the standard
hints coming from Coq and Chargu\'eraud's TLC
library\footnote{\url{http://www.chargueraud.org/softs/tlc/}}, which is
used both by JSCert and our development).

\section{Conclusion}

In the course of our work, we found numerous cases where
\lambdajs{} did not faithfully follow the ES5 
specification\footnote{\url{https://github.com/tilk/LambdaCert/wiki/Issues-with-LambdaS5}}.
Some of these issues were serious enough that
far-reaching changes in the \lambdajs{} environment were needed
to solve them. Interestingly, no serious issues in JSCert were
found, which strengthens our confidence that it accurately models
the ES5 specification.

While proving the correctness of while loops, we found an interesting
inconsistency between the ES5 specification and the behavior
of the two most popular JavaScript implementations, Google's V5 and
Mozilla's SpiderMonkey. The discovery of the inconsistency,
which involves exceptions and statement result values,
resulted in a correction being made in the official ES6
specification\footnote{\url{https://esdiscuss.org/topic/loop-unrolling-and-completion-values-in-es6}}.

The proof scripts we developed are very taxing for the Coq
proof assistant -- the compilation of the whole development
requires several gigabytes of RAM memory, and takes several
hours to compile, even though efforts were made to optimize it.
We are still looking for a solution for the performance issues.

The development is still not complete, but it already
includes most of the important core features of the
JavaScript language. We are looking forward to receiving
feedback from the community.

%\appendix
%\section{Appendix Title}

%This is the text of the appendix, if you need one.

%\acks

%Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

\bibliography{mybib}

% The bibliography should be embedded for final submission.

%\begin{thebibliography}{}
%\softraggedright

%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%P. Q. Smith, and X. Y. Jones. ...reference text...

%\end{thebibliography}

\end{document}

%                       Revision History
%                       -------- -------
%  Date         Person  Ver.    Change
%  ----         ------  ----    ------

%  2013.06.29   TU      0.1--4  comments on permission/copyright notices

### 2016-12-12 paper/paper.tex
adding notes
itp style

@@
-%-----------------------------------------------------------------------------
-%
-%               Template for sigplanconf LaTeX Class
-%
-% Name:         sigplanconf-template.tex
-%
-% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
-%               file for SIGPLAN conference proceedings.
-%
-% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
-%               sigplanconf-guide.pdf
-%
-% Author:       Paul C. Anagnostopoulos
-%               Windfall Software
-%               978 371-2316
-%               paul@windfall.com
-%
-% Created:      15 February 2005
-%
-%-----------------------------------------------------------------------------
-
-\documentclass{sigplanconf}
-
-% The following \documentclass options may be useful:
-
-% preprint      Remove this option only once the paper is in final form.
-% 10pt          To set in 10-point type instead of 9-point.
-% 11pt          To set in 11-point type instead of 9-point.
-% authoryear    To obtain author/year citation style instead of numeric.
+\documentclass{llncs}
 
@@
 
+\usepackage{natbib}
+
 \urlstyle{sf}
@@
 
-\special{papersize=8.5in,11in}
-\setlength{\pdfpageheight}{\paperheight}
-\setlength{\pdfpagewidth}{\paperwidth}
+\title{Certified Desugaring of Javascript Programs using Coq}
 
-\conferenceinfo{COQPL '16}{January 23, 2015, St. Petersburg, Florida, United States} 
-\copyrightyear{2016} 
-\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
-\doi{nnnnnnn.nnnnnnn}
-
-% Uncomment one of the following two, if you are not going for the 
-% traditional copyright transfer agreement.
-
-%\exclusivelicense                % ACM gets exclusive license to publish, 
-                                  % you retain copyright
-
-\permissiontopublish             % ACM gets nonexclusive license to publish
-                                  % (paid open-access papers, 
-                                  % short abstracts)
-
-\titlebanner{banner above paper title}        % These are ignored unless
-\preprintfooter{short description of paper}   % 'preprint' option specified.
-
-\title{Certified Desugaring of Javascript Programs using Coq}
-\subtitle{Experience Report}
-
-\authorinfo{Marek Materzok}
-           {INRIA, University of Wroc\l{}aw}
-           {marek.materzok@inria.fr}
+\author{Marek Materzok\inst{1} \and Alan Schmitt\inst{2}}
+\institute{University of Wroc\l{}aw, Poland\\\email{marek.materzok@ii.uni.wroc.pl}
+\and Inria, France\\\email{alan.schmitt@inria.fr}}
 
@@
 
-\category{D.3.1}{Programming Languages}{Formal Definitions and Theory}
+\section{Notes}
 
-% general terms are not compulsory anymore, 
-% you may leave them out
-%\terms
-%term1, term2
-
-\keywords
-Coq, JavaScript, ECMAScript, mechanized semantics
+\begin{itemize}
+\item give figures on test coverage
+\end{itemize}
 
@@
 
-%\appendix
-%\section{Appendix Title}
-
-%This is the text of the appendix, if you need one.
-
-%\acks
-
-%Acknowledgments, if needed.
-
-% We recommend abbrvnat bibliography style.
-
 \bibliographystyle{abbrvnat}
@@
 
-% The bibliography should be embedded for final submission.
-
-%\begin{thebibliography}{}
-%\softraggedright
-
-%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
-%P. Q. Smith, and X. Y. Jones. ...reference text...
-
-%\end{thebibliography}
-
 \end{document}
-
-%                       Revision History
-%                       -------- -------
-%  Date         Person  Ver.    Change
-%  ----         ------  ----    ------
-
-%  2013.06.29   TU      0.1--4  comments on permission/copyright notices

### 2016-12-12 LambdaS5/tests/jsparser.sh
Updates for the latest jscert.

@@ -5,5 +5,5 @@ BASE_PATH=`dirname $0`/..
 cd $BASE_PATH/tests
 
 #NODE_PATH=/usr/lib/node_modules/ exec node $BASE_PATH/tests/jsparser_node.js $1
-exec js jsparser_js.js esprima.js $1
+exec js24 jsparser_js.js esprima.js $1
 

### 2016-12-12 paper/paper.tex

@@
 
+\usepackage{todonotes}
+\newcommand{\as}[1]{\todo[color=green]{#1}}
+\newcommand{\asi}[1]{\todo[color=green,inline]{#1}}
+
+\usepackage{xspace}
+
 \begin{document}
@@
 
-\newcommand{\lambdajs}{$\lambda_\textrm{JS}$}
+\newcommand{\lambdajs}{$\lambda_\textrm{JS}$\xspace}
 
@@
 A number of semantics were developed for the JavaScript language.
-We are specifically interested in two of them: JSCert and \lambdajs{}.
+We are specifically interested in two of them: JSCert and \lambdajs.
 In order to increase our confidence that the two semantics correctly
@@
 \item give figures on test coverage
+\item say we created tests for bugs
 \end{itemize}
@@
 ECMAScript. Several versions of the standard now exist, the latest
-two being ECMAScript 5.1 (ES5, released in 2011) and ECMAScript 6
-(ES6, released in 2015). Its role is to make portability between different
+ones being ECMAScript 5.1 (ES5, released in 2011), ECMAScript 2015, and
+ECMAScript 2016.\footnote{The ECMAScript standard is now published yearly.} Its role is to make portability between different
 JavaScript implementations possible by specifying the intended behavior
@@
 
-One such semantics is JSCert~\cite{Bodin-al:POPL14}, which has the
-goal of formalizing the ECMAScript (version 5.1) specification in the form of
-a pretty-big-step semantics~\cite{Chargueraud:ESOP13} expressed in 
+One such semantics is \lambdajs. It aims to capture
+the semantics of JavaScript by defining a simple core language
+(based on the untyped call-by-value lambda calculus), and then
+transforming (or ``desugaring'') the full JavaScript language into this
+simple core. The \lambdajs semantics was originally developed to model the 
+ECMAScript 3 specification~\citep{Guha-al:ECOOP10}, and it was later
+revised~\citep{Politz-al:DLS12} to handle ECMAScript 5.1
+features.\footnote{The authors named the revised language S5; however,
+we continue to name it \lambdajs for simplicity.}
+The interpreter for the core language and the JavaScript-to-core
+transformation were developed using 
+OCaml;\footnote{\url{https://github.com/brownplt/LambdaS5}}
+the interpreter was tested using the official test262 test suite.
+The \lambdajs semantics is useful for tool developers, as it
+allows the tools to analyze JavaScript programs by working
+on the simple core language of \lambdajs, having all the complexities
+of JavaScript handled in the desugared code. It was already
+successfully used to verify AdSafe. \as{add citation}
+
+More recently, JSCert~\citep{Bodin-al:POPL14},
+formalized the ECMAScript 5.1 specification in the form of
+a pretty-big-step semantics~\citep{Chargueraud:ESOP13} expressed in 
 the form of a Coq inductive predicate. It aims to be as close to the
@@
 
-Another semantics we are interested in is \lambdajs. It aims to capture
-the semantics of JavaScript by defining a simple core language
-(based on the untyped call-by-value lambda calculus), and then
-transforming (or ``desugaring'') the full JavaScript language into this
-simple core. The \lambdajs{} semantics was originally developed to model the 
-ECMAScript 3 specification~\cite{Guha-al:ECOOP10}, and it was later
-revised~\cite{Politz-al:DLS12} to handle ECMAScript 5.1
-features\footnote{The authors named the revised language S5; however,
-we continue to name it \lambdajs{} for simplicity.}.
-The interpreter for the core language and the JavaScript-to-core
-transformation were developed using 
-OCaml\footnote{\url{https://github.com/brownplt/LambdaS5}};
-the interpreter was tested using the official test262 test suite.
-The \lambdajs{} semantics could be useful for tool developers, as it
-allows the tools to analyze JavaScript programs by working
-on the simple core language of \lambdajs, having all the complexities
-of JavaScript handled in the desugared code. It was already
-successfully used to verify AdSafe. %TODO citation
-
 We believe is worthwhile to formally relate the two semantics. One reason is,
@@
 and is likely to find inconsistencies or other problems. Another reason is
-that having \lambdajs{} formalized and certified (by proving correctness
+that having \lambdajs formalized and certified (by proving correctness
 with respects to JSCert) enables development of certified analysis tools
-using \lambdajs{} as the core.
+using \lambdajs as the core.
 
@@
 to the complexities of the JavaScript language. The ES5 specification,
-upon which JSCert and \lambdajs{} are based, is 258 pages long and 
+upon which JSCert and \lambdajs are based, is 258 pages long and 
 has 16 chapters. The JSCert pretty-big-step semantics, which models
 the specification, has over 900 rules; the entire Coq source code
-for the semantics has over 10000 lines. The \lambdajs{} environment
+for the semantics has over 10000 lines. The \lambdajs environment
 file, which contains the ECMAScript algorithms expressed in core
-\lambdajs{} code, is almost 5000 lines long.
+\lambdajs code, is almost 5000 lines long.
 
 To reduce the complexity of the task, we are currently working
-only on proving \lambdajs{} correct with respect to JSCert,
-i.e., that if the \lambdajs{} program successfully terminates,
+only on proving \lambdajs correct with respect to JSCert,
+i.e., that if the \lambdajs program successfully terminates,
 then its result matches the result given by JSCert. We are also
@@
 correctness of non-strict features only when it is convenient.
+\asi{How difficult would it be to prove the other direction (if we have a
+  derivation in JSCert, then we have one in \lambdajs?)}
 
@@
 
-\section{Formalization of \lambdajs{}}
+\section{Formalization of \lambdajs}
 
 For the purpose of relating the two semantics, we first
-developed a formalization of the \lambdajs{} core language.
+developed a formalization of the \lambdajs core language.
 This was necessary because no such formalization existed previously;
-a formalization of the original, ES3-based \lambdajs{}
-was available\footnote{\url{https://github.com/brownplt/LambdaJS}},
-but the two versions of \lambdajs{} differ too much for this
+a formalization of the original, ES3-based \lambdajs
+was available,\footnote{\url{https://github.com/brownplt/LambdaJS}}
+but the two versions of \lambdajs differ too much for this
 to be useful.
 
-We decided to formalize the \lambdajs{} core using a
+We decided to formalize the \lambdajs core using a
 pretty-big-step semantics, as in JSCert, to make the job of
 relating the two semantics easier. We have also developed
-an interpreter for \lambdajs{}, which we proved sound and
+an interpreter for \lambdajs, which we proved sound and
 complete with respect to the pretty-big-step semantics.
@@
 to be very slow, and also generated large proof terms, which increased
-the memory footprint. The problem is well-known~\cite{Monin:Coq2}; 
+the memory footprint. The problem is well-known~\citep{Monin:Coq2}; 
 our solution involved
@@
 We have also implemented in Coq the other two crucial components
-of the \lambdajs{} semantics: the desugaring function and the 
+of the \lambdajs semantics: the desugaring function and the 
 environment. The desugaring was implemented as Coq functions,
 taking JavaScript statements and expressions (as defined in
-JSCert) to \lambdajs{} expressions. The environment, which
+JSCert) to \lambdajs expressions. The environment, which
 implements the complexities of the JavaScript semantics
-in \lambdajs{} code, was translated from \lambdajs{} syntax
+in \lambdajs code, was translated from \lambdajs syntax
 to Coq definitions with the help of an OCaml parser and
@@
 
-\section{Correctness proof for \lambdajs{} desugaring}
+\section{Correctness proof for \lambdajs desugaring}
 
@@
 a set of predicates specifying the relationship between 
-JavaScript and \lambdajs{} language features. As \lambdajs{} was
+JavaScript and \lambdajs language features. As \lambdajs was
 designed to implement JavaScript, the relationship is fairly
 simple and natural. One place where the two languages differ
-is the object heap, which is used in \lambdajs{} to implement
+is the object heap, which is used in \lambdajs to implement
 several JavaScript language features, including the
@@
 relation, which relates JavaScript objects and environment records
-to \lambdajs{} objects.
+to \lambdajs objects.
 
@@
 which state that, e.g., the JavaScript objects and their
-\lambdajs{} representations are correctly formed. 
+\lambdajs representations are correctly formed. 
 The invariants are split into two relations, where one
@@
 The correctness proof is structured using well-founded 
-induction on the depth of the \lambdajs{} pretty-big-step
+induction on the depth of the \lambdajs pretty-big-step
 semantics derivation. This grants a lot of freedom in structuring
 the proof, which is important because of certain structural
-differences between JSCert and \lambdajs{}, which involve,
+differences between JSCert and \lambdajs, which involve,
 e.g., handling of ES5 references. We use three kinds of induction
@@
 In the course of our work, we found numerous cases where
-\lambdajs{} did not faithfully follow the ES5 
+\lambdajs did not faithfully follow the ES5 
 specification\footnote{\url{https://github.com/tilk/LambdaCert/wiki/Issues-with-LambdaS5}}.
 Some of these issues were serious enough that
-far-reaching changes in the \lambdajs{} environment were needed
+far-reaching changes in the \lambdajs environment were needed
 to solve them. Interestingly, no serious issues in JSCert were
@@
 
-\bibliographystyle{abbrvnat}
+\bibliographystyle{plain}
 

### 2016-12-12 paper/paper.tex
looking at the things to show from the Coq

@@
 \usepackage{hyperref}
-\usepackage{listings,lstlangcoq}
 
@@
 \usepackage{xspace}
+
+\usepackage{minted}
 
@@
 
-For the purpose of relating the two semantics, we first
-developed a formalization of the \lambdajs core language.
-This was necessary because no such formalization existed previously;
-a formalization of the original, ES3-based \lambdajs
-was available,\footnote{\url{https://github.com/brownplt/LambdaJS}}
-but the two versions of \lambdajs differ too much for this
-to be useful.
+For the purpose of relating the two semantics, we first developed a
+formalization of the \lambdajs core language. This was necessary because no such
+formalization existed previously; a formalization of the original, ES3-based
+\lambdajs was available,\as{I propose we use biblatex and add bib references for
+  urls instead of
+  footnotes.}\footnote{\url{https://github.com/brownplt/LambdaJS}} but the two
+versions of \lambdajs differ too much for this to be useful.
 
@@
 which made it simple and robust.
+
+\begin{minted}{coq}
+Inductive resultof (T : Type) : Type :=
+| result_some : T -> resultof T
+| result_fail : string -> resultof T        (* program error *)
+| result_impossible : string -> resultof T  (* inconsistency *)
+| result_bottom : resultof T                (* out of fuel *)
+| result_dump : ctx -> store -> resultof T  (* dump state *)
+
+Definition eval runs c st (e : expr) : result :=
+  let return_value := result_value st in
+  match e with
+  | expr_empty => return_value value_empty
+  | expr_undefined => return_value value_undefined
+...
+
+Record runs_type : Type := runs_type_intro {
+    runs_type_eval : ctx -> store -> expr -> result
+}.
+
+Definition runs_0 := {|
+    runs_type_eval := fun c st _ => result_bottom
+  |}
+.
+
+Definition runs_S runs := 
+  let wrap {A : Type} (f : runs_type -> ctx -> store -> A) c st : A :=
+    f runs c st
+  in {|
+    runs_type_eval := wrap eval
+  |}
+.
+
+Fixpoint runs max_step : runs_type :=
+  match max_step with
+  | 0 => runs_0
+  | S max_step' => runs_S (runs max_step')
+  end
+.
+
+Lemma eval_complete : forall c st e o,
+  red_expr c st e o -> 
+  exists k, runs_type_eval (runs k) c st e = result_some o.
+
+Record runs_type_correct runs :=
+    make_runs_type_correct {
+        runs_type_correct_eval : forall c st e o,
+            runs_type_eval runs c st e = result_some o -> red_expr c st e o
+}.
+
+Lemma eval_correct : forall runs c st e o,
+  runs_type_correct runs -> 
+  eval runs c st e = result_some o ->
+  red_expr c st e o.
+
+Lemma runs_correct : forall k, runs_type_correct (runs k). 
+\end{minted}
+
+\asi{Add the theorem.}
 
@@
 
+\asi{Put a forward pointer to a section where we described the issue in more
+  details.}
+
 We have also implemented in Coq the other two crucial components
@@
 and takes almost a megabyte of disk space!
+
+\asi{we should put a code snippet of the environment here. Also, do we want to
+  talk more about the OCaml tool used to generate it?}
 
@@
 
-For the proof, we also need to keep track of certain invariants,
-which state that, e.g., the JavaScript objects and their
-\lambdajs representations are correctly formed. 
-The invariants are split into two relations, where one
-involves the evaluation contexts, and the other -- the heaps.
-Both of these involve the bisimulation relation, but the
-context invariant uses it only positively. As the bisimulation
-relation only grows during the evaluation (as new objects
-are created), this makes it possible to keep the proofs simple.
+For the proof, we also need to keep track of certain invariants, which state
+that, e.g., the JavaScript objects and their \lambdajs representations are
+correctly formed. The invariants are split into two relations, where one
+involves the evaluation contexts, and the other the heaps. Both of these involve
+the bisimulation relation, but the context invariant uses it only positively. As
+the bisimulation relation only grows during the evaluation (as new objects are
+created but objects are never collected), this makes it possible to keep the
+proofs simple.
 
@@
 hints coming from Coq and Chargu\'eraud's TLC
-library\footnote{\url{http://www.chargueraud.org/softs/tlc/}}, which is
+library,\footnote{\url{http://www.chargueraud.org/softs/tlc/}} which is
 used both by JSCert and our development).
@@
 \lambdajs did not faithfully follow the ES5 
-specification\footnote{\url{https://github.com/tilk/LambdaCert/wiki/Issues-with-LambdaS5}}.
+specification.\footnote{\url{https://github.com/tilk/LambdaCert/wiki/Issues-with-LambdaS5}}
 Some of these issues were serious enough that
@@
 resulted in a correction being made in the official ES6
-specification\footnote{\url{https://esdiscuss.org/topic/loop-unrolling-and-completion-values-in-es6}}.
+specification.\footnote{\url{https://esdiscuss.org/topic/loop-unrolling-and-completion-values-in-es6}}
+\asi{We should detail the bug.}
 
@@
 \end{document}
+
+%%% Local Variables:
+%%% coding: utf-8
+%%% TeX-engine: xetex_sh
+%%% End:

### 2016-12-13 paper/paper.tex
additional notes on what to say
Added reference to the ADsafe paper.
Converted to biblatex and biber.

@@
 
-\usepackage{natbib}
+\usepackage[backend=biber,bibencoding=utf8]{biblatex}
+
+\addbibresource{mybib.bib}
 
@@
 \newcommand{\asi}[1]{\todo[color=green,inline]{#1}}
+\newcommand{\mm}[1]{\todo[color=blue]{#1}}
+\newcommand{\mmi}[1]{\todo[color=blue,inline]{#1}}
 
@@
 \item say we created tests for bugs
+\item say thanks to Valentin Lorentz, to the people at Brown
+\item related work: KJS
 \end{itemize}
@@
 simple core. The \lambdajs semantics was originally developed to model the 
-ECMAScript 3 specification~\citep{Guha-al:ECOOP10}, and it was later
-revised~\citep{Politz-al:DLS12} to handle ECMAScript 5.1
+ECMAScript 3 specification~\cite{Guha-al:ECOOP10}, and it was later
+revised~\cite{Politz-al:DLS12} to handle ECMAScript 5.1
 features.\footnote{The authors named the revised language S5; however,
@@
 of JavaScript handled in the desugared code. It was already
-successfully used to verify AdSafe. \as{add citation}
+successfully used to verify ADsafe~\cite{Politz-al:SEC11}.
 
-More recently, JSCert~\citep{Bodin-al:POPL14},
+More recently, JSCert~\cite{Bodin-al:POPL14},
 formalized the ECMAScript 5.1 specification in the form of
-a pretty-big-step semantics~\citep{Chargueraud:ESOP13} expressed in 
+a pretty-big-step semantics~\cite{Chargueraud:ESOP13} expressed in 
 the form of a Coq inductive predicate. It aims to be as close to the
@@
   derivation in JSCert, then we have one in \lambdajs?)}
+\mmi{It would require inverting the JSCert semantic judgments.
+  Is this working now? I remember Petar trying to do this.}
 
@@
 to be very slow, and also generated large proof terms, which increased
-the memory footprint. The problem is well-known~\citep{Monin:Coq2}; 
+the memory footprint. The problem is well-known~\cite{Monin:Coq2}; 
 our solution involved
@@
 
-\bibliographystyle{plain}
-
-\bibliography{mybib}
+\printbibliography
 

### 2016-12-13 paper/paper.tex
Moved some URLs from footnotes to the bibliography.

@@
 transformation were developed using 
-OCaml;\footnote{\url{https://github.com/brownplt/LambdaS5}}
+OCaml\cite{github:brownplt/LambdaS5};
 the interpreter was tested using the official test262 test suite.
@@
 formalization existed previously; a formalization of the original, ES3-based
-\lambdajs was available,\as{I propose we use biblatex and add bib references for
-  urls instead of
-  footnotes.}\footnote{\url{https://github.com/brownplt/LambdaJS}} but the two
+\lambdajs was available\cite{github:brownplt/LambdaJS}, but the two
 versions of \lambdajs differ too much for this to be useful.
@@
 hints coming from Coq and Chargu\'eraud's TLC
-library,\footnote{\url{http://www.chargueraud.org/softs/tlc/}} which is
+library~\cite{TLC}, which is
 used both by JSCert and our development).
@@
 resulted in a correction being made in the official ES6
-specification.\footnote{\url{https://esdiscuss.org/topic/loop-unrolling-and-completion-values-in-es6}}
+specification\cite{esdiscuss:loop-unrolling}.
 \asi{We should detail the bug.}

### 2016-12-13 paper/paper.tex
ES5 bug explained.

@@
 \newcommand{\mmi}[1]{\todo[color=blue,inline]{#1}}
+
+\newcommand{\jsinline}[1]{\texttt{#1}}
 
@@
 
-\section{Conclusion}
+\section{Issues found}
+
+\subsection{Issues in the \lambdajs definition}
 
@@
 
+\subsection{An issue in the ES5 specification}
+
 While proving the correctness of while loops, we found an interesting
 inconsistency between the ES5 specification and the behavior
-of the two most popular JavaScript implementations, Google's V5 and
-Mozilla's SpiderMonkey. The discovery of the inconsistency,
-which involves exceptions and statement result values,
+of the two most popular JavaScript implementations, Google's V8 and
+Mozilla's SpiderMonkey. The issue involves a subtle interaction
+between the \jsinline{break} statements, looping constructs
+and statement completion values (which can be observed only by \jsinline{eval}).
+ES5 specifies that if the loop body breaks to a~label outside the loop,
+the completion values returned by the loop body in the previous iterations
+are forgotten. For instance, ES5 specifies the following code to return~1 when run by \jsinline{eval}:
+\begin{minted}{javascript}
+var x; 
+l: { 
+    x = 1; 
+    while(true) if (x) x = x - 1; else break l 
+}
+\end{minted}
+This kind of semantics is problematic because it breaks the loop unrolling
+transformation. If we unroll the \jsinline{while} loop once, we get the
+following code, which ES5 specifies to return~0:
+\begin{minted}{javascript}
+var x; 
+l: { 
+    x = 1; 
+    if (true) { 
+        if (x) x = x - 1; else break l; 
+        while(true) if (x) x = x - 1; else break l 
+    }
+}
+\end{minted}
+The two pieces of code, when run by the major JavaScript engines, both return~0.
+The discovery of the inconsistency
 resulted in a correction being made in the official ES6
 specification\cite{esdiscuss:loop-unrolling}.
-\asi{We should detail the bug.}
+
+\subsection{Issues in JSCert}
+
+\mmi{TODO}
+
+\section{Conclusion}
 

### 2016-12-14 paper/paper.tex
tiny comment

@@
 \usepackage{todonotes}
+
+\definecolor{aliceblue}{rgb}{0.94, 0.97, 1.0}
+
 \newcommand{\as}[1]{\todo[color=green]{#1}}
 \newcommand{\asi}[1]{\todo[color=green,inline]{#1}}
-\newcommand{\mm}[1]{\todo[color=blue]{#1}}
-\newcommand{\mmi}[1]{\todo[color=blue,inline]{#1}}
+\newcommand{\mm}[1]{\todo[color=aliceblue]{#1}}
+\newcommand{\mmi}[1]{\todo[color=aliceblue,inline]{#1}}
 
@@
 correctness of non-strict features only when it is convenient.
-\asi{How difficult would it be to prove the other direction (if we have a
-  derivation in JSCert, then we have one in \lambdajs?)}
-\mmi{It would require inverting the JSCert semantic judgments.
-  Is this working now? I remember Petar trying to do this.}
+\asi{We should have a sentence like ``Proving the other direction requires doing
+  an inversion of the full JSCert semantics. Unfortunately, the techniques we
+  have used (see Section $\ldots$) do not scale to semantics of that size.'' }
 

### 2016-12-14 paper/paper.tex
working on the introduction

@@
 
-The JavaScript language is standardized by ECMA under the name
-ECMAScript. Several versions of the standard now exist, the latest
-ones being ECMAScript 5.1 (ES5, released in 2011), ECMAScript 2015, and
-ECMAScript 2016.\footnote{The ECMAScript standard is now published yearly.} Its role is to make portability between different
-JavaScript implementations possible by specifying the intended behavior
-of the various language constructs and built-in functions of JavaScript.
-The specification also makes it possible to develop formal semantics
-for the language, which can then be used for, e.g., developing
-static analysis tools.
+The JavaScript language is standardized by ECMA under the name ECMAScript.
+Several versions of the specification now exist, the latest ones being
+ECMAScript 5.1 (ES5, released in 2011), ECMAScript 2015, and ECMAScript
+2016.\footnote{The ECMAScript standard is now published yearly.} The role of the
+specification is to make portability between different JavaScript
+implementations possible by specifying the intended behavior of the various
+language constructs and built-in functions of JavaScript. The specification also
+makes it possible to develop formal semantics for the language, which can then
+be used to, e.g., develop static analysis tools or guarantee that the invariants
+required by the specification can be satisfied.
 
@@
 transformation were developed using 
-OCaml\cite{github:brownplt/LambdaS5};
+OCaml~\cite{github:brownplt/LambdaS5};
 the interpreter was tested using the official test262 test suite.
@@
 on the simple core language of \lambdajs, having all the complexities
-of JavaScript handled in the desugared code. It was already
+of JavaScript handled in the desugared code. It has already been
 successfully used to verify ADsafe~\cite{Politz-al:SEC11}.
 
-More recently, JSCert~\cite{Bodin-al:POPL14},
-formalized the ECMAScript 5.1 specification in the form of
-a pretty-big-step semantics~\cite{Chargueraud:ESOP13} expressed in 
-the form of a Coq inductive predicate. It aims to be as close to the
-(informal English) specification as possible by maintaining close correspondence
-between the evaluation rules and steps in the ES5 pseudocode. The project
-also includes JSRef, a reference interpreter for JavaScript written in Coq
-and proven correct with respect to JSCert. The interpreter is also tested
-using the official ECMA test suite, test262. Thanks to the correctness proof,
-testing of JSRef increases the confidence we can have that JSCert
-correctly models ES5.
+More recently, JSCert~\cite{Bodin-al:POPL14}, formalized the ECMAScript 5.1
+specification in the form of a pretty-big-step
+semantics~\cite{Chargueraud:ESOP13} expressed in the form of a Coq inductive
+predicate. It aims to be as close to the (informal English) specification as
+possible by maintaining close correspondence between the evaluation rules and
+steps in the ES5 pseudocode. The project also includes JSRef, a reference
+interpreter for JavaScript written in Coq and proven correct with respect to
+JSCert. The interpreter is also tested using the official ECMA test suite,
+test262. Thanks to the correctness proof, testing of JSRef increases the
+confidence that JSCert faithfully models ES5.
 
-We believe is worthwhile to formally relate the two semantics. One reason is,
-because the two semantics use a different approach, relating
-them will increase the confidence we can place in both of them,
-and is likely to find inconsistencies or other problems. Another reason is
-that having \lambdajs formalized and certified (by proving correctness
-with respects to JSCert) enables development of certified analysis tools
-using \lambdajs as the core.
+The goal of this work is to formally relate the two semantics. Our motivation
+is twofold. First, as the two semantics use a different approach, relating them
+increases the confidence we can place in both of them, and has the side effect
+of uncovering inconsistencies or other problems. Second, a formal and certified
+version of \lambdajs, obtained by proving its correctness with respects to
+JSCert, enables development of certified analysis tools using \lambdajs as the
+core.
 
-This is easier said than done. The task is very complex, mostly due
-to the complexities of the JavaScript language. The ES5 specification,
-upon which JSCert and \lambdajs are based, is 258 pages long and 
-has 16 chapters. The JSCert pretty-big-step semantics, which models
-the specification, has over 900 rules; the entire Coq source code
-for the semantics has over 10000 lines. The \lambdajs environment
-file, which contains the ECMAScript algorithms expressed in core
-\lambdajs code, is almost 5000 lines long.
+The main challenge lies in the complexities of the JavaScript language. The ES5
+specification, upon which JSCert and \lambdajs are based, is 258 pages long and
+has 16 chapters. The JSCert pretty-big-step semantics, which models the
+specification, has over 900 rules; the entire Coq source code for the semantics
+has over 10000 lines. The \lambdajs environment file, which contains the
+ECMAScript algorithms expressed in core \lambdajs code, is almost 5000 lines
+long.
+
+As is done in \lambdajs, we focus on the \emph{strict mode} restriction of
+JavaScript. Our contributions are as follows.
+\begin{enumerate}
+\item Two formalizations in Coq of \lambdajs: an inductive one, and an
+  executable one. The formalizations coincide: they are shown to be correct and
+  complete in relation to each other. This contribution is described in
+  Section~\ref{sec:form-lambdajs}.
+\item A formalization in Coq of desugaring from JavaScript to \lambdajs, as well
+  as the initial \lambdajs environment in Coq. This contribution is described in
+  Section~\ref{sec:desugaring}.
+\item A correctness proof of the desugaring, stating that if a JavaScript
+  program translated to \lambdajs successfully terminates, then the initial
+  program has a semantics according to JSCert and the results match. This is
+  described in Section~\ref{sec:corr-proof-lambdajs}.
+\end{enumerate}
+
+Note that we do not prove the converse, i.e., that if a JavaScript program has a
+semantic in JSCert, then its \lambdajs translation successfully terminates and
+the results match. The reason is technical and is described in
+Section~\ref{sec:coq-techniques} alongside the techniques we implemented in Coq
+for this development.
 
@@
 \section{Formalization of \lambdajs}
+\label{sec:form-lambdajs}
+
+\asi{describe the changes to the language.}
 
@@
 
+\section{Desugaring}
+\label{sec:desugaring}
+
 We have also implemented in Coq the other two crucial components
@@
 \section{Correctness proof for \lambdajs desugaring}
+\label{sec:corr-proof-lambdajs}
 
@@
 used both by JSCert and our development).
+
+\section{Coq techniques}
+\label{sec:coq-techniques}
+
+Proving the other direction requires doing an inversion of the full JSCert
+semantics. Unfortunately, the techniques we have used do not scale to semantics
+of that size.
 

### 2016-12-14 paper/paper.tex
more progress on the introduction

@@
 
-\usepackage[backend=biber,bibencoding=utf8]{biblatex}
+\usepackage[backend=biber,bibencoding=utf8,urldate=iso8601,date=iso8601]{biblatex}
 
@@
 semantic in JSCert, then its \lambdajs translation successfully terminates and
-the results match. The reason is technical and is described in
-Section~\ref{sec:coq-techniques} alongside the techniques we implemented in Coq
-for this development.
+the results match. For such a result, we would need to apply \textsf{inversion}
+on the full JSCert semantics, which does not currently work. This issue is
+described in Section~\ref{sec:coq-techniques} alongside the techniques we
+implemented in Coq for this development.
 
-To reduce the complexity of the task, we are currently working
-only on proving \lambdajs correct with respect to JSCert,
-i.e., that if the \lambdajs program successfully terminates,
-then its result matches the result given by JSCert. We are also
-restricting our attention to ECMAScript strict mode, proving the
-correctness of non-strict features only when it is convenient.
-\asi{We should have a sentence like ``Proving the other direction requires doing
-  an inversion of the full JSCert semantics. Unfortunately, the techniques we
-  have used (see Section $\ldots$) do not scale to semantics of that size.'' }
-
-We were able to prove correctness of a large part of the
-JavaScript semantics, including function calls, lexical environment
-handling, and most operations involving objects. The work is still
-in progress to prove the remaining parts, which include integer
-operations, the
-defineOwnProperty method, and elements of the standard library.
-
-The development is publicly available and can be downloaded from GitHub:
-
-\url{https://github.com/tilk/LambdaCert}
+The development is publicly available and can be downloaded from
+GitHub~\cite{github:lambdacert}.
 
@@
 \label{sec:corr-proof-lambdajs}
+
+\asi{This was in the introduction, we need to integrate it again.
+
+We were able to prove correctness of a large part of the
+JavaScript semantics, including function calls, lexical environment
+handling, and most operations involving objects. The work is still
+in progress to prove the remaining parts, which include integer
+operations, the
+defineOwnProperty method, and elements of the standard library.
+}
 

### 2016-12-14 paper/paper.tex
Added subsection on precomputed inversion.

@@
 
-At this stage we stumbled upon an interesting performance issue 
-involving the Coq proof assistant. The proof of completeness required
-inverting the pretty-big-step semantics judgment; this turned out
-to be very slow, and also generated large proof terms, which increased
-the memory footprint. The problem is well-known~\cite{Monin:Coq2}; 
-our solution involved
-precomputing the required inversion principles using
-\texttt{Derive Inversion}, and using the precomputed inversion
-principles to create a specialized inversion tactic.
+% This was moved.
+%At this stage we stumbled upon an interesting performance issue 
+%involving the Coq proof assistant. The proof of completeness required
+%inverting the pretty-big-step semantics judgment; this turned out
+%to be very slow, and also generated large proof terms, which increased
+%the memory footprint. The problem is well-known~\cite{Monin:Coq2}; 
+5our solution involved
+%precomputing the required inversion principles using
+%\texttt{Derive Inversion}, and using the precomputed inversion
+%principles to create a specialized inversion tactic.
 
-\asi{Put a forward pointer to a section where we described the issue in more
-  details.}
+%\asi{Put a forward pointer to a section where we described the issue in more
+%  details.}
 
@@
 
+\subsection{Precomputed inversion}
+
+Developing the correctness proof for the JavaScript desugaring required
+inverting the \lambdajs pretty-big-step semantics judgment; this
+turned out to be very slow, and also generated large proof terms, which 
+increased the memory footprint. The problem is well-known~\cite{Monin:Coq2}.
+To solve it, we precomputed the required inversion principles
+using \texttt{Derive Inversion}, e.g.:
+\begin{minted}{coq}
+Derive Inversion inv_red_exprh_if with (forall k c st e e1 e2 oo,
+    red_exprh k c st (expr_if e e1 e2) oo) Sort Prop.
+\end{minted}
+Using these, we created a specialized inversion tactic
+which selects and applies the required induction principle.
+The \texttt{Derive Inversion} calls and the tactic definitions
+are generated automatically from the Coq definition of the semantics.
+
+Unfortunately, this technique was still not efficient enough to be
+applied to JSCert: the definition of the JavaScript semantics in
+JSCert has around 10 times more semantic rules than the definition
+in \lambdajs.
+
+\subsection{Proof automation}
+
 \section{Issues found}

### 2016-12-14 LambdaS5/tests/jsparser.sh
JSCert's parser now used.

@@ -1,9 +1,0 @@
-#!/bin/bash
-
-BASE_PATH=`dirname $0`/..
-
-cd $BASE_PATH/tests
-
-#NODE_PATH=/usr/lib/node_modules/ exec node $BASE_PATH/tests/jsparser_node.js $1
-exec js24 jsparser_js.js esprima.js $1
-

### 2016-12-15 paper/paper.tex
Simpler correctness/completeness.

@@
 | result_bottom : resultof T                (* out of fuel *)
-| result_dump : ctx -> store -> resultof T  (* dump state *)
+| result_dump : ctx -> store -> resultof T. (* dump state *)
 
-Definition eval runs c st (e : expr) : result :=
+Definition eval_S eval c st (e : expr) : result :=
   let return_value := result_value st in
@@
 
-Record runs_type : Type := runs_type_intro {
-    runs_type_eval : ctx -> store -> expr -> result
-}.
+Definition eval_0 : eval_fun := fun c st _ => result_bottom.
 
-Definition runs_0 := {|
-    runs_type_eval := fun c st _ => result_bottom
-  |}
-.
-
-Definition runs_S runs := 
-  let wrap {A : Type} (f : runs_type -> ctx -> store -> A) c st : A :=
-    f runs c st
-  in {|
-    runs_type_eval := wrap eval
-  |}
-.
-
-Fixpoint runs max_step : runs_type :=
+Fixpoint eval_k max_step : eval_fun :=
   match max_step with
-  | 0 => runs_0
-  | S max_step' => runs_S (runs max_step')
+  | 0 => eval_0
+  | S max_step' => eval_S (eval_k max_step')
   end
@@
 
-Lemma eval_complete : forall c st e o,
+Lemma eval_k_complete : forall c st e o,
   red_expr c st e o -> 
-  exists k, runs_type_eval (runs k) c st e = result_some o.
+  exists k, eval_k k c st e = result_some o.
 
-Record runs_type_correct runs :=
-    make_runs_type_correct {
-        runs_type_correct_eval : forall c st e o,
-            runs_type_eval runs c st e = result_some o -> red_expr c st e o
-}.
+Record eval_fun_correct eval := forall c st e o,
+  eval c st e = result_some o -> red_expr c st e o.
 
-Lemma eval_correct : forall runs c st e o,
-  runs_type_correct runs -> 
-  eval runs c st e = result_some o ->
-  red_expr c st e o.
-
-Lemma runs_correct : forall k, runs_type_correct (runs k). 
+Lemma eval_k_correct : forall k, runs_type_correct (eval k). 
 \end{minted}

### 2016-12-15 paper/paper.tex
Some changes to lambdajs described.

@@
 
-\asi{describe the changes to the language.}
+Because of the various issues we found in \lambdajs as defined
+in~\cite{Politz-al:DLS12} and which we detail in Section~\ref{sec:issues},
+we decided to revise the language. The full list of changes is
+detailed in~TODO\mm{Can we just link the wiki page? Or should we use an appendix?}, we present some of them below.
+\begin{itemize}
+\item The attribute access and modification operators, which tried to directly
+    model the operators from JavaScript, were removed. We found that the complexities
+    of attribute access is much better handled in the desugaring.
+\item We removed the imperative features of the language which do not involve the
+    object heap. This change simplified the definition of the semantics, which made
+    the proofs easier.
+\item We added a new primitive value named \texttt{empty}, which models the empty
+    completion value of JavaScript, and a second kind of the semicolon operator,
+    which returns the last non-empty value of its subexpressions. This allowed to
+    correctly implement statement completion values in a natural way.
+\item An integer datatype was added. Even though ES5 itself has no integers,
+    they are used in the definition of the binary operators (bitwise logical operators,
+    bit shifting).
+\end{itemize}
 
-For the purpose of relating the two semantics, we first developed a
+For the purpose of relating the semantics of \lambdajs and ES5, we developed a
 formalization of the \lambdajs core language. This was necessary because no such
@@
 an interpreter for \lambdajs, which we proved sound and
-complete with respect to the pretty-big-step semantics.
+complete with respect to the pretty-big-step 
+semantics.\footnote{The initial version of the interpreter was
+written by Martin Lorentz.}
 The proof script was developed using domain-specific tactics,
@@
 %the memory footprint. The problem is well-known~\cite{Monin:Coq2}; 
-5our solution involved
+%our solution involved
 %precomputing the required inversion principles using
@@
 \section{Issues found}
+\label{sec:issues}
 

### 2017-01-25 paper/paper.tex
small structural changes

@@
 
-\newcommand{\as}[1]{\todo[color=green]{#1}}
-\newcommand{\asi}[1]{\todo[color=green,inline]{#1}}
+\newcommand{\as}[1]{\todo[color=green!20]{#1}}
+\newcommand{\asi}[1]{\todo[color=green!20,inline]{#1}}
 \newcommand{\mm}[1]{\todo[color=aliceblue]{#1}}
@@
 In order to increase our confidence that the two semantics correctly
-model JavaScript, we try to relate them formally using Coq. The size and complexity
+model JavaScript, we relate them formally using Coq. The size and complexity
 of the two semantics makes this a complex problem with many obstacles
@@
 we decided to revise the language. The full list of changes is
-detailed in~TODO\mm{Can we just link the wiki page? Or should we use an appendix?}, we present some of them below.
+detailed in~TODO\mm{Can we just link the wiki page? Or should we use an
+  appendix?}\as{Let's do both: put them in the appendix, and link to the wiki.}, we present some of them below.
 \begin{itemize}
@@
 The proof script was developed using domain-specific tactics,
-which made it simple and robust.
+which made it simple and robust. They are detailed in
+Section~\ref{subsec:ljstactics}.
+
+\subsection{Coq Formalization}
+\label{subsec:ljscoq}
+
+\asi{starting from the main lemmas, describe the Coq formalization, describing
+  necessary details as they occur.}
 
@@
 
-% This was moved.
-%At this stage we stumbled upon an interesting performance issue 
-%involving the Coq proof assistant. The proof of completeness required
-%inverting the pretty-big-step semantics judgment; this turned out
-%to be very slow, and also generated large proof terms, which increased
-%the memory footprint. The problem is well-known~\cite{Monin:Coq2}; 
-%our solution involved
-%precomputing the required inversion principles using
-%\texttt{Derive Inversion}, and using the precomputed inversion
-%principles to create a specialized inversion tactic.
+\subsection{Tactics}
+\label{subsec:ljstactics}
 
-%\asi{Put a forward pointer to a section where we described the issue in more
-%  details.}
+\asi{Describe the tactics used in the \lambdajs proofs.}
 
